---
title: "Prediction Assignment Writeup"
author: "Mooc Coursera"
date: "December 27, 2015"
output: html_document
---
# Executive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

# Data
The training data for this project are available here:

   *  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
   
The test data are available here:

   * https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

# Preprocessing Step
```{r}
library(caret)
library(Hmisc)
```

```{r}
training<-read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!"))
testing<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!"))
```

Variables with no missing values will be good candidate for Predictor
```{r}
isAnyMissing <- sapply(testing, function (x) any(is.na(x) | x == ""))
isPredictor <- !isAnyMissing & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(isAnyMissing))
predictorCandidates <- names(isAnyMissing)[isPredictor]
predictorCandidates
```
# Subsetting the primary dataset to include only the predictor candidates and the outcome variable,
```{r}
varincluded <- c("classe", predictorCandidates)
training <- training[,varincluded]
dim(training)
```
```{r}
names(training)
```
```{r}
training$classe = factor(training$classe)
```
# Splitting the dataset into 60 percent training and 40 percent test set
